Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-01-16 PAPER.tex
minor sentence change
Slight tweaks to language re: sensitivity

Just as important as the eigenvalues are the eigenvectors of our empirical covariances. In general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the stronger eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we shall see in Section \acl{insert}, this minimizes signal loss. In contrast, the weaker eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. %\acl{Think about whether we need to do the Fourier space version as well.}

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$ K, yielding $T_{sys} = 431$ K at $150$ MHz. 
This is lower than the $T_{sys}$ of $500$ K used in A15 because of several small miscalculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after squaring visibility measurements. In A15, a factor of $2$ was mistakenly used.
\item $N_{lst}$: The number of LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lst} = 8$.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al., \textit{in prep}
use three separation types.
\item $t_{int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In A15, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In A15, this number was 
the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, for the A15 analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}

>>>>

Just as important as the eigenvalues are the eigenvectors of our empirical covariances. In general, different eigenvectors converge to their true forms at different rates. This is illustrated by Figure \ref{fig:toy_sigloss17}, which shows the convergence of eigenvectors in an empirical estimate of a covariance matrix whose true form is a diagonal matrix with eigenvalues spanning four orders of magnitude. We use as a convergence metric $\varepsilon(\widehat{\textbf{v}})$ for the empirical eigenvectors $\widehat{\textbf{v}}$, defined as:
\begin{equation}
\label{eq:converge_eig}
\varepsilon (\widehat{\textbf{v}}) \equiv \sqrt{\sum_{i}^{N_{f}}|\textbf{v}-\widehat{\textbf{v}}|_{i}^2},
\end{equation}
where $N_{f}$ is the number of frequencies ($20$) in the mock data. The eigenmode convergence curves  in Figure \ref{fig:toy_sigloss17} are ranked ordered by eigenvalue, such that ``Eigenmode \#0" illustrates the convergence of the eigenvector with the largest eigenvalue, ``Eigenmode \#1" for the second largest eigenvalue, and so on. One sees that the stronger eigenmodes converge to their true eigenvectors more quickly. With only a small number of realizations, these empirically estimated modes already retain little correlation with the specific realizations of data that were used to form the empirical covariance. As we will see in the next section, using only the strongest eigenmodes, which are less coupled to data realizations, minimizes signal loss. In contrast, the weaker eigenmodes retain more memory of the data realizations, which leads to correlations that induce signal loss. %\acl{Think about whether we need to do the Fourier space version as well.}

@@

where $\nu$ are frequencies in GHz. We use a receiver temperature of $144$ K, yielding $T_{sys} = 431$ K at $150$ MHz. 
This is lower than the $T_{sys}$ of $500$ K used in A15 because of several small miscalculation errors that were 
identified\footnote{For example, there was a missing a square root in going from a variance to a standard deviation.}.
\item $\sqrt{2}$: This factor in the denominator of the sensitivity equation comes from taking the real part of the power spectrum 
estimates after cross-multiplying independent `even' and `odd' visibility measurements (which is done principally to avoid a noise bias). In A15, a factor of $2$ was mistakenly used.
\item $N_{lst}$: The number of LST bins that go into a power spectrum estimation. The sensitivity scales as the square root 
because we integrate incoherently over time. For PAPER-64, $N_{lst} = 8$.
\item $N_{seps}$: The number of baseline separation types averaged incoherently in a final power spectrum estimate. For the 
analysis in this paper, we only use one type of baseline, hence $N_{seps}=1$. The updated limits in Kolopanis et al., \textit{in prep}
use three separation types.
\item $t_{int}$: Length of an independent integration of the data. It is crucial to adapt this number if filtering is applied along the time axis (i.e. a 
fringe-rate filter). We compute the effective integration time of our fringe-rate filtered data by scaling the original integration time $t_{i}$
using the following:
\begin{equation}
t_{int} = t_{i} \frac{\int1 \, df}{\int w^{2}(f) \,df},
\end{equation}
where $t_{i}=43$ seconds, $t_{int}$ is the fringe-rate filtered integration time, $w$ is the fringe-rate profile, and the integral is 
taken over all fringe-rates. For PAPER-64, this number is $t_{int} = 3857$ s. 
\item $N_{days}$: The total number of days of data analyzed. In A15, this number was set to $135$. However, because we 
divide our data in half (to form `even' and `odd' datasets), this number should be reduced by a factor of $2$. Additionally, 
because our LST coverage is not $100\%$ complete (it doesn't overlap for every single day), we compute a more realistic value of 
$N_{days}$ as:
\begin{equation}
 \frac{1}{N_{days}} = \sqrt{\Big\langle\frac{1}{N_{i}^{2}} \Big\rangle_{i}},
 \end{equation}
\noindent where $i$ is over LST (\citealt{jacobs_et_al2015}). For PAPER-64, our revised estimate of $N_{days}$ is $\sim34$ 
days.
\item $N_{bls}$: The number of baselines contributing to the sensitivity of a power spectrum estimate. In A15, this number was 
the total number of $30$ m East/West baselines used in the analysis. However, using the total number of baselines neglects 
the fact that baselines are averaged into groups (see Section \ref{sec:Boot}) for computational speed-up when cross-multiplying data. Our revised estimate for the parameter is:
\begin{equation}
N_{bls} = \frac{N_{bls}}{N_{gps}}\sqrt{N_{gps}^{2}-N_{gps}},
\end{equation}
\noindent where, for the A15 analysis, $N_{gps} = 5$. Each baseline group averages down linearly as the number of baselines 
entering the group ($N_{bls}/N_{gps}$) and then as the square root of the number of cross-multiplied pairs ($\sqrt{N_{gps}^{2} - 
N_{gps}}$). For the revised PAPER-64 analysis with only one baseline separation type, this becomes $N_{bls} \sim 46$ instead 
of $51$. 
\item $N_{pols}$: The number of polarizations averaged together. For the case of Stokes I, $N_{pols}=2$.
\end{itemize}
